Potential Factors of Decreasing Birth Rates in the US since the Great Recession and Beyond

### 2024-02-08 outputs/paper/paper.qmd
Add r scripts for the figure

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

>>>>

@@

## Source

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: fig-usbrate-graph
#| fig-cap: "Trend in US Birth Rates"

# Load necessary libraries
library(ggplot2)

# Set seed for reproducibility
set.seed(123)

# Read the data from CSV file
data <- read.csv(file = here::here("outputs/data/fig_1.csv"))

# Plotting the data
ggplot(data, aes(x = year, y = brate_all)) +
  geom_line(color = "blue") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  annotate("text", x = 2007, y = 80, label = "2007", vjust = -0.5, hjust = 0.5, color = "black") +
  geom_hline(yintercept = seq(50, 80, by = 5), color = "lightgrey") +
  scale_y_continuous(limits = c(50, 80), breaks = seq(50, 80, by = 5)) +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 1), labels = function(x) ifelse(x %% 5 == 0, x, "")) +
  labs(x = "Year", y = "Births per 1,000 women age 15-44",
       title = "Trend in US Birth Rates",
       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
  theme_classic()

```
```{r}
# Load necessary libraries
library(ggplot2)
library(gridExtra)

# Set seed for reproducibility
set.seed(123)

# Read the data from CSV file
data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))

# Plotting the first line graph (Graph A)
graph_A <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
  scale_color_manual(values = c("Age 15-19" = "blue", "Age 20-24" = "orange", "Age 25-29" = "grey",
                                 "Age 30-34" = "yellow", "Age 35-39" = "cyan", "Age 40-44" = "green"),
                     labels = c("Age 15-19", "Age 20-24", "Age 25-29", "Age 30-34", "Age 35-39", "Age 40-44"),
                     name = "Age Group") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5)) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
       title = "Trend in US Birth Rates by Age Group",
       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
  theme_classic()

# Plotting the second line graph (Graph B)
graph_B <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
  scale_color_manual(values = c("Hispanic" = "grey", "Black, non-Hispanic" = "orange", "White, non-Hispanic" = "blue"),
                     labels = c("Hispanic", "Black, non-Hispanic", "White, non-Hispanic"),
                     name = "Race/Ethnicity") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5)) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(x = "Year", y = "Births per 1,000 women in relevant population subgroup",
       title = "Trend in US Birth Rates by Race/Ethnicity",
       caption = "Birth Rates collected from CDC Vital Statistics Births Reports for 2015, 2019, and 2020.") +
  theme_classic()

# Arrange both graphs side by side
grid.arrange(graph_A, graph_B, nrow = 1)

```

```{r}
# Load necessary libraries
library(ggplot2)
library(gridExtra)

# Set seed for reproducibility
set.seed(123)

# Read the data from CSV file
data <- read.csv(file = here::here("outputs/data/figs_2a_2b.csv"))

# Plotting the first line graph (Graph A)
graph_A <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_1519, color = "Age 15-19"), size = 1.2) +
  geom_line(aes(y = brate_2024, color = "Age 20-24"), size = 1.2) +
  geom_line(aes(y = brate_2529, color = "Age 25-29"), size = 1.2) +
  geom_line(aes(y = brate_3034, color = "Age 30-34"), size = 1.2) +
  geom_line(aes(y = brate_3539, color = "Age 35-39"), size = 1.2) +
  geom_line(aes(y = brate_4044, color = "Age 40-44"), size = 1.2) +
  geom_text(aes(label = "Age 15-19", x = 2020, y = brate_1519), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
  geom_text(aes(label = "Age 20-24", x = 2020, y = brate_2024), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
  geom_text(aes(label = "Age 25-29", x = 2020, y = brate_2529), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
  geom_text(aes(label = "Age 30-34", x = 2020, y = brate_3034), hjust = 0, vjust = -0.5, color = "yellow", family = "Times") +
  geom_text(aes(label = "Age 35-39", x = 2020, y = brate_3539), hjust = 0, vjust = -0.5, color = "cyan", family = "Times") +
  geom_text(aes(label = "Age 40-44", x = 2020, y = brate_4044), hjust = 0, vjust = -0.5, color = "green", family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1980, 2020), breaks = seq(1980, 2020, by = 5), labels = function(x) paste(x, "\n")) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(y = "Births per 1,000 women in relevant population subgroup",
       title = "A: Five-year age group",
       caption = "Note: Birth rates by age group, race and ethnicity, and marital status are gathered from CDC Vital Statistics Births Reports. Birt rates by Hispanic subpopulation, mother's education are calculated by the original author's NCHS Vital Statistics birth microdata, SEER population data, and the Current Population Survey.") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"))

# Plotting the second line graph (Graph B)
graph_B <- ggplot(data, aes(x = year)) +
  geom_line(aes(y = brate_hisp, color = "Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_blacknh, color = "Black, non-Hispanic"), size = 1.2) +
  geom_line(aes(y = brate_whitenh, color = "White, non-Hispanic"), size = 1.2) +
  geom_text(aes(label = "Hispanic", x = 2020, y = brate_hisp), hjust = 0, vjust = -0.5, color = "grey", family = "Times") +
  geom_text(aes(label = "Black, non-Hispanic", x = 2020, y = brate_blacknh), hjust = 0, vjust = -0.5, color = "orange", family = "Times") +
  geom_text(aes(label = "White, non-Hispanic", x = 2020, y = brate_whitenh), hjust = 0, vjust = -0.5, color = "blue", family = "Times") +
  geom_vline(xintercept = 2007, linetype = "dotted") +
  geom_hline(yintercept = seq(0, 140, by = 20), color = "lightgrey") +
  scale_x_continuous(limits = c(1990, 2020), breaks = seq(1990, 2020, by = 5), labels = function(x) paste(x, "\n")) +
  scale_y_continuous(limits = c(0, 140), breaks = seq(0, 140, by = 20)) +
  labs(title = "B: Race and ethnicity (ages 15-44)") +
  theme_classic()

# Arrange both graphs side by side
grid.arrange(graph_A, graph_B, nrow = 1)

```

## Methodology

## Features

### Data Cleaning

### Data Modification

### Data Visualization

# Results
The results section should convey findings.
## Table, graph, table, graph
